Wrattler: \textnormal{Reproducible, live and polyglot notebooks}

### 2018-03-27 tapp/paper.tex
Adding Vizer reference

\begin{figure}
\includegraphics[scale=0.15]{screen.png}
\caption{\small{TheGamma script that downloads and aggregates UK government data, running in 
  Wrattler notebook with a live preview.}}
\label{fig:proto}
\end{figure}

@@

Dataflow notebooks \cite{dataflow} are more sophisticated and attach unique hashes to cell 
evaluations. This allows the user to refer to dependencies explicitly and, in effect, construct
a dependecy graph similar to ours manually.
Scientific workflow systems \cite{taverna,kepler} manage evaluation of workflows over a 
dependency graph similarly to Wrattler, but often allow editing it directly via a GUI, rather
than through code in a notebook.
The noWorkflow project \cite{noworkflow} links the two approaches by instrumenting Jupyter 
kernel with a mechanism for capturing provenance based on light-weight code annotations.

Our method of dependency grpah construction is inspired by Roslyn \cite{roslyn} and extends our
earlier work on TheGamma \cite{thegamma}. It is simiar to methods used in live programming 
languages \cite{live,subtext}, as well as incremental compilation \cite{incremental} and
partial evaluation \cite{partial}. Wrattler adapts those methods to a notebook-like data
science environment.

@@

\begin{acks}
We thank to our colleagues from the AIDA (Artificial Intelligence for data analytics) project
Chris Williams, Zoubin Ghahramani and Ian Horrocks and also numerous attendees of a recent 
AIDA workshop in The Alan Turing Institute. This work was in part supported by The Alan Turing 
Institute under the EPSRC grant EP/N510129/1.
\end{acks}

>>>>

\begin{figure}
\vspace{-0.5em}
\includegraphics[scale=0.15]{screen.png}
\caption{\small{TheGamma script that downloads and aggregates UK government data, running in 
  Wrattler notebook with a live preview.}}
\label{fig:proto}
\vspace{-0.5em}
\end{figure}

@@

Dataflow notebooks \cite{dataflow} are attach unique hashes to cell 
evaluations. This allows the user to refer to dependencies explicitly and, in effect, construct
a dependecy graph similar to ours manually.
Scientific workflow systems \cite{taverna,kepler} manage evaluation of workflows over a 
dependency graph similarly to Wrattler, but often allow editing it directly via a GUI, rather
than through code in a notebook.

The noWorkflow project \cite{noworkflow} links the two approaches by instrumenting Jupyter 
kernel with a mechanism for capturing provenance based on light-weight annotations.
Vizer \cite{vizer} focuses on integrating notebooks with spreadsheet-like interface.
It internally uses a data store component similar to ours, but does not keep dependency graph
on the client.

Our dependency grpah construction is inspired by Roslyn \cite{roslyn} and extends an
earlier work on TheGamma \cite{livegamma}. It is simiar to methods used in live programming 
languages \cite{live,subtext}, incremental compilation \cite{incremental} and
partial evaluation \cite{partial}. Wrattler adapts those methods to a notebook environment.

@@

\begin{acks}
We thank to our colleagues from the AIDA (Artificial Intelligence for data analytics) project
Chris Williams, Zoubin Ghahramani and Ian Horrocks and attendees of a recent 
AIDA workshop. This work was in part supported by The Alan Turing 
Institute under the EPSRC grant EP/N510129/1.
\end{acks}
